---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "The Covid Pandemic: Our World in Data"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Nadia Matulich 19880812"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Department of Economics, SU" # First Author's Affiliation
Email1: "nadiamatulich\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: FALSE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
#toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
 This document reports on the Covid pandemic 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.


```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

# Question 1: Covid

Having been directed by editor of the New York Times, the following needs to be done:

* African experience of covid, vs rest of the world 
* Asssess co-morbidities on a Macro level
    * Did countries with higher rates of co-morbidities fare worse?
* Hospitalisation:
     * How quickly did different regions increase their hospitalisation facilities
     * As well as whether this lagged or led ICU emissions

```{r Loading Data: Covid}

deaths_data <- read.csv("data/Covid/Deaths_by_cause.csv")

covid_data <- read.csv("data/Covid/owid-covid-data.csv")

```

If we compare different regions we obtain the following:



```{r Comparing Deaths and Cases by Regions}
#Comparing Deaths and Cases by Regions

regions_compare<-function(vjust_input, size_input, round_input){

    deaths_by_region <- covid_data %>%
        filter(date==last(date)) %>%
        filter(continent==c("Africa","Europe","South America","North America", "Oceania", "Asia")) %>% #sorry I couldn't think of a more elegant way to get rid of the world averages OWID included with blank (But not NA?) continents
        group_by(continent) %>%
        summarise(total_deaths=sum(x=total_deaths_per_million, na.rm=T))

    cases_by_region <- covid_data %>%
        filter(date==last(date)) %>%
        filter(continent==c("Africa","Europe","South America","North America", "Oceania", "Asia")) %>%
        group_by(continent) %>%
        summarise(total_cases=sum(x=total_cases_per_million, na.rm=T))

    cases_and_deaths<-left_join(deaths_by_region, cases_by_region, by='continent')

    g<- cases_and_deaths %>%
        ggplot(aes(x=reorder(continent,total_deaths,decreasing=TRUE), y=total_cases, fill=factor(ifelse(continent=="Africa", "Africa","Other Regions"))))+
        geom_bar(stat="identity")+
        geom_text(aes(label=round(total_deaths,round_input)),vjust=vjust_input, size=size_input)+
        theme_bw()+
        theme(legend.position="none")+
        labs(title="Total Cases Per Million",
             subtitle="with Total Deaths Per Million",
             x="Region",
             y="Total",)+
        scale_fill_brewer(palette="Purples")
    g

}


regions_compare_2<-function(){

    deaths_by_region <- covid_data %>%
        filter(date==last(date)) %>%
        filter(continent==c("Africa","Europe","South America","North America", "Oceania", "Asia")) %>%
        group_by(continent) %>%
        summarise(total_deaths=sum(x=total_deaths_per_million, na.rm=T))

    cases_by_region <- covid_data %>%
        filter(date==last(date)) %>%
        filter(continent==c("Africa","Europe","South America","North America", "Oceania", "Asia")) %>%
        group_by(continent) %>%
        summarise(total_cases=sum(x=total_cases_per_million, na.rm=T))

    cases_and_deaths<-left_join(deaths_by_region, cases_by_region, by='continent')

    cases_and_deaths<-cases_and_deaths %>% mutate(., new_var=total_deaths/total_cases)

    h<- cases_and_deaths %>%
        ggplot(aes(x=reorder(continent,new_var,decreasing=TRUE), y=new_var, fill=factor(ifelse(continent=="Africa", "Africa","Other Regions"))))+
        geom_bar(stat="identity")+
        theme_bw()+
        theme(legend.position="none")+
        labs(title="Proportion of Deaths",
             x="Region",
             y="Total",)+
        scale_fill_brewer(palette="Purples")
    h
}

```

```{r}
regions_compare(-0.5,3,0)
```


```{r}
regions_compare_2()
```



* Africa experienced the highest death rates (per 1mil population) but managed to contain the spread of covid quite well. 
* Europe experienced the highest incidence of cases but exhibited a much lower rate of deaths per 1000000 population.


At the start of the pandemic, a lot was said about the effect of co-morbidities on mortality. 

```{r}

deaths_by_cardio<-function(){

    #need to filter by last date and country
    #severity of covid measured by deaths per population

    deaths_by_country <- covid_data %>%
        filter(date==last(date)) %>%
        group_by(location) %>%
        summarise(deaths_per_1000000=total_deaths/(population/100000))

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, aged_70_older,gdp_per_capita,extreme_poverty,cardiovasc_death_rate,diabetes_prevalence,female_smokers, male_smokers, life_expectancy, human_development_index))

    #cannot feasibly show all of these in 15 hours, so for now consider only co-morbidity data

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, cardiovasc_death_rate,diabetes_prevalence,male_smokers,female_smokers))

    morbidity_deaths<-left_join(deaths_by_country,co_morbidities, by='location')

    i<- morbidity_deaths %>%
        ggplot(aes(x=cardiovasc_death_rate, y=deaths_per_1000000))+
        geom_point()+
        geom_smooth()+
        theme_bw()+
        labs(title="Deaths per 1000000 Population",
             subtitle = "By Cardio Vascular Death Prevalence",
             x="Cardio Vascular Death Prevalence",
             y="Deaths")
    i

}



#sorry this is really inelegant, did not know how to make a function print multiple outputs so in the interest of time I am doing it the inelegant way
#the first wrangling chunk is the same in each of them

deaths_by_diabetes<-function(){

    #need to filter by last date and country
    #severity of covid measured by deaths per population

    deaths_by_country <- covid_data %>%
        filter(date==last(date)) %>%
        group_by(location) %>%
        summarise(deaths_per_1000000=total_deaths/(population/100000))

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, aged_70_older,gdp_per_capita,extreme_poverty,cardiovasc_death_rate,diabetes_prevalence,female_smokers, male_smokers, life_expectancy, human_development_index))

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, cardiovasc_death_rate,diabetes_prevalence,male_smokers,female_smokers))

    morbidity_deaths<-left_join(deaths_by_country,co_morbidities, by='location')

    j<- morbidity_deaths %>%
        ggplot(aes(x=diabetes_prevalence, y=deaths_per_1000000))+
        geom_point()+
        geom_smooth()+
        theme_bw()+
        labs(title="Deaths per 1000000 Population",
             subtitle = "By Diabetes Prevalence",
             x="Diabetes Prevalence",
             y="Deaths")
    j
}



deaths_by_smoking<-function(){

    #need to filter by last date and country
    #severity of covid measured by deaths per population

    deaths_by_country <- covid_data %>%
        filter(date==last(date)) %>%
        group_by(location) %>%
        summarise(deaths_per_1000000=total_deaths/(population/100000))

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, aged_70_older,gdp_per_capita,extreme_poverty,cardiovasc_death_rate,diabetes_prevalence,female_smokers, male_smokers, life_expectancy, human_development_index))

    #cannot feasibly show all of these in 15 hours, so for now consider only co-morbidity data

    co_morbidities <- covid_data %>%
        filter(date==last(date)) %>%
        select(c(location, cardiovasc_death_rate,diabetes_prevalence,male_smokers,female_smokers))

    morbidity_deaths<-left_join(deaths_by_country,co_morbidities, by='location')


    k<- morbidity_deaths %>%
        ggplot(aes(y=deaths_per_1000000))+
        geom_point(aes(x=male_smokers), color="blue")+
        geom_point(aes(x=female_smokers),colour="pink")+
        geom_smooth(aes(x=female_smokers), colour="pink")+
        geom_smooth(aes(x=male_smokers))+
        theme_bw()+
        labs(title="Deaths per 1000000 Population",
             subtitle = "By Smoking Prevalence",
             x="Smoking Prevalence",
             y="Deaths")
    k
}

deaths_by_smoking()
```


```{r Deaths By Co Morbidities}

deaths_by_cardio()

```


```{r Deaths By Co Morbidities 2}

deaths_by_diabetes()

```


```{r Deaths By Co Morbidities 3}

deaths_by_smoking()


```




we see that :
* in countries with low cardio vascular death prevalence, there are high covid deaths 
    * this is likely due to high cardio death prevalence being linked to a poor hospital system with poor record keeping
    * as cardio vascular deaths prevalence increases, deaths decrease, then increase again (then decrease again)
     * initial decrease likely due to better record keeping
     * subsequent increase due to underlying co-morbities
     * subsequent decrease due to lack of data at the end of the graph (hence high confidence intervals)


