---
title: "Question 4"
author: "Nadia Matulich"
date: "`r Sys.Date()`"
output: pdf_document
---

````{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
	
)

library(pacman)
pacman::p_load(devtools, tinytex, rmarkdown,tidyverse, xts, dplyr, tibble, lubridate, glue)



```
#Question 4
 
 (using a regular markdown for speed as time is running out and it was not specified in the question)
Question 4


```{r Loading Data: Netflix}

netflix_titles<-read.csv("./data/netflix/titles.csv")

netflix_credits<-read.csv("./data/netflix/credits.csv")

netflix<-left_join(netflix_credits,netflix_titles, by="id")

```

Would like to investigate most credited directors


```{r Most Popular director}

most_credited<-function(movie_or_series, actor_or_director){
    
most_popular_directors<-netflix %>% filter(release_year>=2010, type==movie_or_series, role==actor_or_director) %>% filter(production_countries=="['US']") %>% count(name) %>% arrange(desc(n))

head(most_popular_directors, 11)

}

table<-most_credited(movie_or_series="MOVIE",actor_or_director="DIRECTOR")

```

next, would like to consider if longer movies are better- 

```{r}

does_runtime_matter<-function(){
    
g<- netflix_titles %>% 
    ggplot(aes(x=runtime,y=imdb_score, colors=type))+
    geom_point(aes(size=imdb_score), alpha=0.1, color="dark blue")+
    geom_smooth()+
    labs(title="Does Run Time Matter?",
         x="Run Time",
         y="IMDB Score")+
    facet_wrap(~type)+
    theme(legend.position="none")+
    theme_bw()
    
g

}

does_runtime_matter()

```

It appears that longer run times are associated with better movies

Lastly, would like to check if older movies are better, and whether things have actually gone downhill with the 'on demad' revolution

```{r}
older_movies_better<-function(){
h<- netflix_titles %>% 
    filter(type=="MOVIE") %>% 
    ggplot(aes(x=release_year, y=imdb_score))+
    geom_point()+
    geom_smooth()+
    theme_bw()+
    labs(title="Were Movies Better in the 90s?",
         subtitle="Or are we just nostalgic?",
         x="Run Time",
         y="IMDB Score")
        
    
h
}
older_movies_better()
```
It appears that they are slightly better

(sorry, really ran out of time towards the end)



